---
title: "Simple Regression Analysis"
author: "AW"
date: "2022/7/10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Key terms

* Dependent vaiable
* Expanatory variable, independent variable, regressor
* Disturbance term
* Residual


## 1. The simple linear model

**The model**:

$$Y_{i}=\beta_{1}+\beta_{2}X_{i}+\mu_{i}$$
where $Y$ is usually described as the **dependent variable**, and $X$ as the **explanatory variable** or **independent variable** or the **regressor**. In statistical analysis, one generally acknowledgesthe fact that the relationship is not exact by explicitly including in it a random factor known as the **disturbance term**, $\mu_{i}$.

**The fitted model**:

$$\hat{Y_{i}}=\hat{\beta_{1}}+\hat{\beta_{2}}X_{i}$$

The difference between the actual value of $Y_{i}$ and the fitted value $\hat{Y_{i}}, in observation $i$ is known as the **residual** in the observation $i$. It will be denoted \hat{\mu_{i}}:

$$\hat{\mu_{i}}=Y_{i}-\hat{Y_{i}}$$




```{r}
# Example in page 98 of the text
eawe21 <- read.csv('./Data/EAWE21.csv', header=T)   # Import data
linear_model <- lm(EARNINGS ~ S, data = eawe21)     # Run the model
linear_model                                        # Print the result
```


Please refer to [Dougherty][Dougherty] and the [book][book].

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

[Dougherty]: https://www.econometrics-with-r.org/index.html
[book]: D:/OneDrive/Bookshelf/Introduction to Econometrics_Dougherty_5th Ed
